Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2022-10-31 ba_resubmit/kundinger_resubmit.tex
JErrys edits

\begin{aug}
\author{\fnms{Brian} \snm{Kundinger}\thanksref{addr1}\ead[label=e1]{brian.kundinger@duke.edu}},
\author{\fnms{Jerome} \snm{Reiter}\thanksref{addr1}\ead[label=e2]{jreiter@duke.edu}}
\and
\author{\fnms{Rebecca C.} \snm{Steorts}\thanksref{addr2}\ead[label=e3]{beka@stat.duke.edu}}

@@

\begin{abstract}
Recently, researchers have developed Bayesian versions of the Fellegi Sunter model for record linkage. These have the crucial advantage of quantifying uncertainty from imperfect linkages.  However, current implementations of Bayesian Fellegi Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks. To address these computational considerations, we propose fast beta linkage (\texttt{fabl}), an extension to the Beta Record Linkage (\texttt{BRL}) method of \cite{sadinle_bayesian_2017}. Specifically, we propose independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead. This also allows us to complete pairwise record comparisons over large datasets through parallel computing and to reduce memory costs through a new technique called storage efficient indexing. Through simulations and two case studies, we show that \texttt{fabl} has markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}.
%
%\textcolor{blue}{We propose a variation on Bayesian Fellegi Sunter models that we call fast beta linkage, or \texttt{fabl}, and is motivated by computational considerations.  Our method extends that of \cite{sadinle_bayesian_2017}, who extended the early work of \cite{fellegi_theory_1969}.}

@@

%% ** Keywords **
\begin{keyword}%[class=MSC]
\kwd{entity resolution}
\kwd{data fusion}
\kwd{hashing}
\kwd{record linkage}
\kwd{Bayesian}
\kwd{distributed computing}
\end{keyword}

@@

In this paper, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability. Following the suggestion in \cite{wortman2019}, we relax the one-to-one matching requirement of \texttt{BRL} and propose independent priors over the matching space. This allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large datasets via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing. These contributions allow \texttt{fabl} to perform record linkage on much larger datafiles than previous Bayesian Fellegi Sunter models at significantly increased speed with minimal loss of accuracy. In particular, computation time under \texttt{BRL} grows quadratically, with the size of each datafile, while computation time under \texttt{fabl} grows linearly, only with the size of the smaller datafile.

@@

Consider two datafiles $\bm{X}_1$ and $\bm{X}_2$ containing records $\{x_{1i}\}_{i=1}^{n_1}$ and $\{x_{2j}\}_{j=1}^{n_2}$ respectively. Without loss of generality, denote files such that $n_1 \geq n_2$. We follow the convention established by \cite{sadinle_bayesian_2017} and say ``record i $\in X_1$" rather than the more compact $x_{1i}$ in order to avoid double subscripts. In the context of bipartite matching, we assume that there are duplications across, but not within, each datafile. Under this framework, the set of matches across datafiles can be represented in two equivalent ways. First, we may use a matrix $\Delta \in \{0, 1\}^{n_1 \times n_2}$, where
\begin{align}
	\Delta_{ij} =
	\begin{cases}
		1, & \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		0, & \text{otherwise}.\\
	\end{cases}
\end{align}
Though intuitive, this sparse matrix representation can become cumbersome for large linkage tasks. More compactly, bipartite matching also can be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in datafile $\bm{X}_2$ such that 
\begin{align}
	Z_{j} =
	\begin{cases}
		i, & \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		n_1 + j, & \text{if record}\;  j \in \bm{X}_2 \; \text{does not have a match in datafile}\; \bm{X}_1. \\
	\end{cases}
\end{align}
We can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot) = 1$ when the expression inside the parentheses is true, and $I(\cdot) = 0$ otherwise.

Denote the set of matches by $\bm{M} = \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 1\}$, and the set of non-matches by 
$\bm{U} =  \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 0\}.$ The record linkage task can be viewed as identifying the sets of  $\bm{M}$ and  $\bm{U}.$ We refer to record pairs that are estimated as matches as ``links'' and to record pairs that are estimated as non-matches as ``non-links.''

@@

Intuitively, matching records (those that refer to the same entity) should be similar; records that are not non-matching should not be similar. \cite{fellegi_theory_1969} proposed encoding this is using a comparison vector $\bm{\gamma}_{ij}$ computed for each record pair $(i,j)$ in $\bm{X}_1 \times \bm{X}_2.$ Denote the number of criteria for comparing records by $F$, such that $\bm{\gamma}_{ij} = (\gamma_{ij}^1, \gamma_{ij}^2, \ldots, \gamma_{ij}^f, \ldots, \gamma_{ij}^F).$ In most cases, $\bm{\gamma}_{ij}$ consists of one comparison for each feature shared between the two datasets. We refer to set of all comparison vectors $\bm{\gamma}_{ij}$ as $\Gamma$.

The simplest way to compare two records is to check for agreement or disagreement, and this is commonly used for categorical variables. For numerical data, we can use absolute difference, and for text data, we can use string distance metrics such as Levenstein or Jaro-Winkler distance \citep{cohen2003comparison}. We can then set thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}. Let $\mathcal{S}_f(i,j)$ denote a general similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f +1$ intervals denoted by $I_{f0}, I_{f1}, \ldots, I_{fL_f}$. Following convention, $I_{f0}$ represents the highest level of agreement (inclusive of complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement). Thus, we can construct comparison vectors in the following way: $\gamma_{ij}^f = l \; \text{if} \; \mathcal{S}_f(i,j) \in I_{fl}.$ The choices of $I_{fl}$ are application specific, which we discuss in our simulation and case studies.

@@

\subsection{Fellegi-Sunter Model}
\label{fellegi-sunter}

The Fellegi Sunter model employs two independence assumptions: first, that comparison vectors are conditionally independent given their matching status, and second, that the matching status of the record pairs are independent. This allows us to model the comparison data using mixture models in the following way:
\begin{align}
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \\
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \notag\\
	& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(p), \notag
\end{align}
where $\mathcal{M}$ and $\mathcal{U}$ are the distributions for matching and non-matching record pairs, $\bm{m}$ and $\bm{u}$ are their respective sets of parameters, and $p$ is the marginal probability that a record pair is a match. When using comparison vectors with discrete agreement levels as described above, the $\bm{m}$ parameter takes the form $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = P(\gamma_{ij}^f = l|Z_j = i)$ for all fields $f$ and agreement levels $l$. The $\bm{u}$ parameters are defined similarly, with $u_{fl} = P(\gamma_{ij}^f = l|Z_j \neq i)$. The $\bm{m}$ and $\bm{u}$ parameters are used to calculate the likelihood ratio  
\begin{align}
	\label{eqn:wts}
	w_{ij} = \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}.
\end{align}
This ratio will be large when there is strong evidence of the pair being a match, and small otherwise.

@@

The Fellegi Sunter method then uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as non-links (with those such that $T_u < w_{ij} < T_m$ left undetermined and subject to clerical review). This is a problem in practice as transitive closures can be violated. Since each $w_{ij}$ is calculated independently, there can be situations in which both $w_{ij} > T_m$ and $w_{i'j} > T_m$, so both $(i,j)$ and $(i', j)$ are declared as links, even though such a matching is not allowed by assumption.

To address this issue, \cite{jaro1989} proposed solving the following linear sum assignment problem:
\begin{align}
	\label{eqn:jaro}
	&\max_{\Delta} \sum_{i=1}^{n_1} \sum_{j=1}^{n_2} w_{ij} \Delta_{ij} 
	\quad \text{subject to} \quad \Delta_{ij} \in \{0,1\}; \\
	& \quad \quad \sum_{i=1}^{n_1}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_2; \; \text{and} \notag\\
	& \quad \quad\sum_{i=1}^{n_2}  \Delta_{ij}  \leq 1, i=1,2, \ldots n_1. \notag
\end{align}

The solution to this problem, for which several simple algorithms exist, is a bipartite matching that maximizes the sum of the Fellegi Sunter weights among matched pairs. Jaro provided no theoretical justification for using this approach; however \cite{sadinle_bayesian_2017} recently showed that under certain conditions, (\ref{eqn:jaro}) is the maximum likelihood estimate for bipartite matching.

@@

\subsection{Beta Record Linkage Model}
\label{BRL}

To estimate a one-to-one matching without using a post-processing step, \cite{sadinle_bayesian_2017} incorporated this constraint directly into the model through a prior distribution over the matching space. He proposed the  beta distribution for bipartite matching, given by
$$P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) + \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
$$
where $\text{B}(\cdot, \cdot)$ represents the Beta function, hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$, and $n_{12}(\bm{Z}) = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$ is the number of records in $\bm{X}_2$ that have a match in $\bm{X}_1$. This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate records when sampling \(Z_j\). This creates a dependency that makes the sampler inherently serial, which can be slow when working on linkage tasks with more than a few thousand records.

%One criticism of the independent matching assumption is that the decision rule leads to ``many-to-many assignments,'' breaking one-to-one matching assumptions. One way of resolving this issue to incorporate the constraint directly into the model as follows:
%\begin{align}
%	&\bm{\Gamma}_{ij} \mid Z_j =i \stackrel{iid}{\sim} \mathcal{M}(\bm{m}),  \\
%	&\bm{\Gamma}_{ij} \mid Z_j \neq i \stackrel{iid}{\sim} \mathcal{U}(\bm{u}) \notag, \\
%	& \bm{Z} \sim \mathcal{B}, \notag
%\end{align}
%where $\mathcal{B}$ represents a prior on the space of bipartite matchings, and $\mathcal{M}(\bm{m})$ and $\mathcal{U}(\bm{u})$ are models for the comparison vectors of matches and non-matches. For \texttt{BRL}, \cite{sadinle_bayesian_2017} proposed the beta distribution for bipartite matching, given by
%$$P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) - \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
%$$
%where $\text{B}(\cdot, \cdot)$ represents the Beta function, hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$, and $n_{12}(\bm{Z}) = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$ is the records in $\bm{X}_2$ that have a match in $\bm{X}_1$. This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate records when sampling \(Z_j\). This creates a dependency that makes the sampler inherently serial, which can slow down the sampler when working on large linkage tasks.

@@

In contrast to the prior over the vector $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we follow \cite{wortman2019} and use independent priors for each component $Z_j$. However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, we use a proper prior. We denote the fast beta prior as follows:
$$Z_j | \pi =
\begin{cases} 
	\frac{1}{n_1}\pi,  & z_j \leq n_1; \\
	1-\pi, &  z_j  = n_1 + j; \\
\end{cases}$$
$$\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}).$$
This set of priors says that record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\pi$, and that each record $i \in \bm{X}_1$ is equally likely to be that match. The parameter $\pi$ is given hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ that encode prior beliefs about  the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}.$

% I don't really like using remarks. If its not relevante enough for the main body, we can just remove. 
%\begin{remark}
%One non-informative choice might be \(\pi \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that non-matches and matches are equally likely, and another might be \(\pi \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\bm{Z}\). Note the prior from \cite{wortman2019} can be viewed as a special case of the fabl prior, with $\pi$ fixed at the mean of a $\text{Beta}\left(1, \frac{1}{n_1}\right)$ random variable. 
%\end{remark}

@@

Importantly, this independence relaxes the one-to-one requirement from \texttt{BRL}; the Gibbs sampler under \texttt{fabl} does ensure that each record in \(\bm{X}_2\) can be matched to at most one record in \(\bm{X}_1\), but allows for the possibility that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\). To resolve many-to-one matchings and obtain a one-to-one estimate, we use the post processing procedure explained in Section \ref{bayes-estimate}.

Lastly, we note that it is common to encounter missing information in datasets used for record linkage. Whether information is missing in record $i \in \bm{X}_1$ or $j \in \bm{X}_2$, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values. We handle this issue by assuming that this missingness occurs at random (MAR, by \cite{LittleRubin2002}). With this assumption, we can marginalize over the missing data, and do all computation simply using the observed data. To notate this, we use the indicator $I_{obs}(\cdot)$, which takes on the value 1 when the argument inside the parentheses is observed, and 0 otherwise.

>>>>

\begin{aug}
\author{\fnms{Brian} \snm{Kundinger}\thanksref{addr1}\ead[label=e1]{brian.kundinger@duke.edu}},
\author{\fnms{Jerome P.} \snm{Reiter}\thanksref{addr1}\ead[label=e2]{jreiter@duke.edu}}
\and
\author{\fnms{Rebecca C.} \snm{Steorts}\thanksref{addr2}\ead[label=e3]{beka@stat.duke.edu}}

@@

\begin{abstract}
%Recently, researchers have developed Bayesian versions of the Fellegi Sunter model for record linkage. These have the crucial advantage of quantifying uncertainty from imperfect linkages.  
\textcolor{red}{Within the field of record linkage, Bayesian methods have the crucial advantage of quantifying uncertainty from imperfect linkages.} However, current implementations of Bayesian Fellegi Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks. To address these computational considerations, we propose fast beta linkage (\texttt{fabl}), an extension to the Beta Record Linkage (\texttt{BRL}) method of \cite{sadinle_bayesian_2017}. Specifically, we use independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead. This also allows us to complete pairwise record comparisons over large datasets through parallel computing and to reduce memory costs through a new technique called storage efficient indexing. Through simulations and two case studies, we show that \texttt{fabl} has markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}.
%
%\textcolor{blue}{We propose a variation on Bayesian Fellegi Sunter models that we call fast beta linkage, or \texttt{fabl}, and is motivated by computational considerations.  Our method extends that of \cite{sadinle_bayesian_2017}, who extended the early work of \cite{fellegi_theory_1969}.}

@@

%% ** Keywords **
\begin{keyword}%[class=MSC]
\kwd{data fusion}
\kwd{distributed computing}
\kwd{entity resolution}
\kwd{hashing}
\kwd{record linkage}
\end{keyword}

@@

In this paper, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability. Following the suggestion in \cite{wortman2019}, we relax the one-to-one matching requirement of \texttt{BRL} and use independent priors over the matching space. This allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large datasets via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing. These contributions allow \texttt{fabl} to perform record linkage on much larger datafiles than previous Bayesian Fellegi Sunter models at significantly increased speed with minimal loss of accuracy. In particular, computation time under \texttt{BRL} grows quadratically, with the size of each datafile, while computation time under \texttt{fabl} grows linearly, only with the size of the smaller datafile.

@@

Consider two datafiles $\bm{X}_1$ and $\bm{X}_2$ containing records $\{x_{1i}\}_{i=1}^{n_1}$ and $\{x_{2j}\}_{j=1}^{n_2}$ respectively. Without loss of generality, denote files such that $n_1 \geq n_2$. We follow the convention established by \cite{sadinle_bayesian_2017} and say ``record $i \in X_1$" rather than the more compact $x_{1i}$ \textcolor{red}{for ease of readability}. Under bipartite matching, the set of matches across datafiles can be represented in two equivalent ways. First, we may use a matrix $\Delta \in \{0, 1\}^{n_1 \times n_2}$, where
\begin{align}
	\Delta_{ij} =
	\begin{cases}
		1, & \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		0, & \text{otherwise}.\\
	\end{cases}
\end{align}
Though intuitive, this sparse matrix representation can become cumbersome for large linkage tasks. More compactly, bipartite matching also can be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in $\bm{X}_2$ such that 
\begin{align}
	Z_{j} =
	\begin{cases}
		i, & \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		n_1 + j, & \text{if record}\;  j \in \bm{X}_2 \; \text{does not have a match in}\; \bm{X}_1. \\
	\end{cases}
\end{align}
We can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot) = 1$ when the expression inside the parentheses is true, and $I(\cdot) = 0$ otherwise.

Denote the set of matches by $\bm{M} = \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 1\}$, and the set of non-matches by 
$\bm{U} =  \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 0\}.$ The record linkage task can be viewed as identifying  $\bm{M}$ and  $\bm{U}.$ We refer to record pairs that are estimated as matches as ``links'' and to record pairs that are estimated as non-matches as ``non-links.''

@@

Intuitively, matching records (those that refer to the same entity) should be similar; records that are non-matching should be dissimilar. \cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\bm{\gamma}_{ij}$ computed for each record pair $(i,j)$ in $\bm{X}_1 \times \bm{X}_2.$ Denote the number of criteria for comparing records by $F$, such that $\bm{\gamma}_{ij} = (\gamma_{ij}^1, \gamma_{ij}^2, \ldots, \gamma_{ij}^f, \ldots, \gamma_{ij}^F).$ In most cases, $\bm{\gamma}_{ij}$ consists of one comparison for each feature shared between the two datasets. We refer to set of all comparison vectors $\bm{\gamma}_{ij}$ as $\Gamma$.

The simplest way to compare any particular feature for two records is to check for exact agreement, and this is commonly used for categorical features. For numerical features, we can use absolute difference between the two feature values. For text features, such as names, we can use string distance metrics such as Levenstein or Jaro-Winkler distance \citep{cohen2003comparison}. We then set thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}.

Let $\mathcal{S}_f(i,j)$ denote a general similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f +1$ intervals denoted by $I_{f0}, I_{f1}, \ldots, I_{fL_f}$. Following convention, $I_{f0}$ represents the highest level of agreement (inclusive of complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement). Thus, we can construct comparison vectors such that $\gamma_{ij}^f = l \; \text{if} \; \mathcal{S}_f(i,j) \in I_{fl}.$ The choices of $I_{fl}$ are application specific, as we discuss in the simulation and case studies.

\textcolor{red}{In the construction of comparison vectors, we note it is common to encounter missing information in datasets used for record linkage. Whether information is missing in record $i \in \bm{X}_1$ or $j \in \bm{X}_2$, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values. We assume that this missingness occurs at random (MAR, by \cite{LittleRubin2002}). With the MAR, we can marginalize over the missing data, and do all computation simply using the observed data. To notate this, we use the indicator $I_{obs}(\cdot)$, which takes on the value 1 when the argument inside the parentheses is observed, and 0 otherwise.}

@@

\subsection{Fellegi-Sunter Models}
\label{fellegi-sunter}

The seminal \cite{fellegi_1969} model employs two independence assumptions: first, that comparison vectors are conditionally independent given their matching status, and second, that the matching status of the record pairs are independent. These allow us to model the comparison data using a mixture model, where
\begin{align}
	\label{eqn:fs_model}
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \\
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \notag\\
	& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(\lambda). \notag
\end{align}
Here $\mathcal{M}$ and $\mathcal{U}$ are the distributions for matching and non-matching record pairs, $\bm{m}$ and $\bm{u}$ are their respective sets of parameters, and $\lambda$ is the marginal probability that a record pair is a match. When using comparison vectors with discrete agreement levels, $\mathcal{M}$ and $\mathcal{U}$ are collections of independent multinomial distributions for each linkage feature. Accordingly, $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = P(\gamma_{ij}^f = l|\Delta_{ij} = 1)$ for all fields $f$ and agreement levels $l$. The $\bm{u}$ parameters are defined similarly, with $u_{fl} = P(\gamma_{ij}^f = l|\Delta_{ij} = 0)$.

\textcolor{red}{Each record pair is then independently classified as a match or nonmatch according to the} likelihood ratio  
\begin{align}
	\label{eqn:wts}
	w_{ij} = \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}.
\end{align}
This ratio is large when there is strong evidence of the pair being a match, and small otherwise. All record pairs with weights above a specified threshold are considered a match. Such independent classifications often leads to matchings that violate one-on-one matching assumptions, which prompted \cite{jaro1989} to formulate an optimization problem that produces a bipartite matching maximizing the sum of the Fellegi Sunter weights among matched pairs.

@@

%The Fellegi Sunter method then uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as non-links (with those such that $T_u < w_{ij} < T_m$ left undetermined and subject to clerical review). This is a problem in practice as transitive closures can be violated. Since each $w_{ij}$ is calculated independently, there can be situations in which both $w_{ij} > T_m$ and $w_{i'j} > T_m$, so both $(i,j)$ and $(i', j)$ are declared as links, even though such a matching is not allowed by assumption.

%To address this issue, \cite{jaro1989} proposed solving the following linear sum assignment problem:
%\begin{align}
%	\label{eqn:jaro}
%	&\max_{\Delta} \sum_{i=1}^{n_1} \sum_{j=1}^{n_2} w_{ij} \Delta_{ij} 
%	\quad \text{subject to} \quad \Delta_{ij} \in \{0,1\}; \\
%	& \quad \quad \sum_{i=1}^{n_1}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_2; \; \text{and} \notag\\
%	& \quad \quad\sum_{i=1}^{n_2}  \Delta_{ij}  \leq 1, i=1,2, \ldots n_1. \notag
%\end{align}
%
%The solution to this problem, for which several simple algorithms exist, is a bipartite matching that maximizes the sum of the Fellegi Sunter weights among matched pairs. Jaro provided no theoretical justification for using this approach; however \cite{sadinle_bayesian_2017} recently showed that under certain conditions, (\ref{eqn:jaro}) is the maximum likelihood estimate for bipartite matching.

@@

%\subsection{Beta Record Linkage Model}
%\label{BRL}

%To estimate a one-to-one matching without using a post-processing step, \cite{sadinle_bayesian_2017} incorporated this constraint directly into the model through a prior distribution over the matching space.

\textcolor{red}{To estimate a one-to-one matching without using a post-processing step, \cite{sadinle_bayesian_2017} incorporated this constraint directly into the model through a Bayesian framework. In addition to standard Dirichlet priors for the $\bm{m}_f$ and $\bm{u}_f$ parameters, he proposed the "beta distribution for bipartite matching" for the linkage parameter $\bm{Z}$. To construct this prior, he assigned a prior to the probability of the indicator that a record in $\bm{X}_2$ had a match in $\bm{X}_1$, so that $I(Z_j \leq n_1) \sim \text{Bernoulli}(\pi)$, where $\pi$ itself is taken to be distributed $\text{Beta}(\alpha_{\pi}, \beta_{\pi})$. It follows that the number of records in $\bm{X}_2$ that have matches, denoted $n_{12}(\bm{Z}) = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$, is distributed according to a $\text{Beta-Binomial}(n_2, \alpha_{\pi}, \beta_{\pi})$. Conditioning on the set of records in $\bm{X}_2$ that have matches, formally denoted $\{I(Z_j \leq n_1)\}_{j=1}^{n_2}$, all $n_1 ! / (n_1 - n_{12}(\bm{Z})!$ bipartite matchings are taken to be equally likely.} Thus the beta distribution for bipartite matching, is given by
\begin{align}
\label{eqn:sadinle_prior}
P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) + \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
\end{align}
where $\text{B}(\cdot, \cdot)$ represents the Beta function. This prior strictly enforces one-to-one matching, inducing a Gibbs sampler that removes previously matched records from the set of candidate records when sampling \(Z_j\). This makes the sampler inherently serial, which can be slow when working on linkage tasks with more than a few thousand records.

%hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X}_2$ that have matches in $\bm{X_1}$, and $n_{12}(\bm{Z}) = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$ is the number of records in $\bm{X}_2$ that have a match in $\bm{X}_1$. 
%One criticism of the independent matching assumption is that the decision rule leads to ``many-to-many assignments,'' breaking one-to-one matching assumptions. One way of resolving this issue to incorporate the constraint directly into the model as follows:
%\begin{align}
%	&\bm{\Gamma}_{ij} \mid Z_j =i \stackrel{iid}{\sim} \mathcal{M}(\bm{m}),  \\
%	&\bm{\Gamma}_{ij} \mid Z_j \neq i \stackrel{iid}{\sim} \mathcal{U}(\bm{u}) \notag, \\
%	& \bm{Z} \sim \mathcal{B}, \notag
%\end{align}
%where $\mathcal{B}$ represents a prior on the space of bipartite matchings, and $\mathcal{M}(\bm{m})$ and $\mathcal{U}(\bm{u})$ are models for the comparison vectors of matches and non-matches. For \texttt{BRL}, \cite{sadinle_bayesian_2017} proposed the beta distribution for bipartite matching, given by
%$$P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) - \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
%$$
%where $\text{B}(\cdot, \cdot)$ represents the Beta function, hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$, and $n_{12}(\bm{Z}) = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$ is the records in $\bm{X}_2$ that have a match in $\bm{X}_1$. This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate records when sampling \(Z_j\). This creates a dependency that makes the sampler inherently serial, which can slow down the sampler when working on large linkage tasks.

@@

In contrast to the prior over the vector $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we follow \cite{wortman2019} and use independent priors for each component $Z_j$. However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, we use a proper prior. We denote the fast Beta prior as follows:
\begin{align}
	\label{eqn:fast_beta_prior}
	Z_j | \pi &= \begin{cases} 
	\frac{1}{n_1}\pi,  & z_j \leq n_1; \\
	1-\pi, &  z_j  = n_1 + j; \\
\end{cases} \\
\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}) \notag.
\end{align}
We can interpret equation \ref{eqn:fast_beta_prior} as follows: record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\pi$, and each record $i \in \bm{X}_1$ is equally likely to be that match. The hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about  the proportion of records in $\bm{X}_2$ that have matches in $\bm{X}_1.$

\textcolor{red}{Previously, \cite{wortman2019} proposed a flat prior over the labeling of $\bm{Z}$, meaning that each value $\{1, \ldots, n_1, n_1 +j\}$ was considered equally likely for $Z_j$. This however amounts to a prior probability of $\frac{n_1}{n_1 + 1}$ that record $j$ has a match in $\bm{X}_1$, and thus tended to over match, leading to poor precision when used in practice. Indeed we find that the flat prior is equivalent to a special case of the fast Beta prior where $\pi$ fixed at the mean of a $\text{Beta}\left(1, \frac{1}{n_1}\right)$ random variable.}

@@

%Importantly, this independence relaxes the one-to-one requirement from \texttt{BRL}; the Gibbs sampler under \texttt{fabl} does ensure that each record in \(\bm{X}_2\) can be matched to at most one record in \(\bm{X}_1\), but allows for the possibility that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\). To resolve many-to-one matchings and obtain a one-to-one estimate, we use the post processing procedure explained in Section \ref{bayes-estimate}.

\textcolor{red}{To obtain an estimate $\hat{\bm{Z}}$ of the linkage structure, we use the loss functions and Bayes estimate from \cite{sadinle_bayesian_2017}, reproduced in Appendix . Since the prior does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$. To achieve a Bayes estimate that fulfills one-to-one matching requirement, we minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$. In the event that we have two records $j$ and $j'$ such that both $P(\hat{Z}_j = i |\Gamma) > \frac{1}{2}$ and $ P(\hat{Z}_{j'} = i |\Gamma) > \frac{1}{2}$, we accept the match with the highest posterior probability, and declare the other to have no match. A similar approach appears in the most probable maximal matching sets used by \cite{steorts_bayesian_2016} to match records to latent entities.}

%It is common to encounter missing information in datasets used for record linkage. Whether information is missing in record $i \in \bm{X}_1$ or $j \in \bm{X}_2$, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values. We assume that this missingness occurs at random (MAR, by \cite{LittleRubin2002}). With the MAR, we can marginalize over the missing data, and do all computation simply using the observed data. To notate this, we use the indicator $I_{obs}(\cdot)$, which takes on the value 1 when the argument inside the parentheses is observed, and 0 otherwise.
